---
title: dplyr_** examples, oh!
author: ~
date: '2017-09-14'
slug: dplyr-example-oh
categories: [r, dplyr, data_munging]
tags: [r, dplyr]
description: 'advanced example on dplyr, from youtube'
---

Today `SuperDataScience` teached an interesting example of using the advanced dplyr functions (summarise_xx, mutate_xx as well as starts_with() and one_of() ) on top of the `AirPassengers` data. 
Watch the video here: https://youtu.be/cavtFdvII4A

And here's the code:

Create the dataset:
```{r}
library(dplyr)
library(tidyverse)

air_passengers <- data.frame(matrix(AirPassengers,
                                    ncol=12,
                                    byrow = T,
                                    dimnames = list(as.character(1:12), month.abb)
                                    )
                             )
air_passengers$year <- as.character(1949:1960)
```

Let's use select() to reorder columns:
```{r}

# reorder column
air_passengers <-  air_passengers %>% 
  select(year, everything())

# what is the max for each variable?
max_flights <- air_passengers %>% 
  summarise_all(max)

# if I want the max for EACH month:
max_each_month <- air_passengers %>% 
    summarise_if(is.numeric, max)

# the chr variable did not get summarised!
str(air_passengers)
```

Here's how to use summarise_at():
```{r}
# mean flights for each month starting with "J"
mean_flights_j <- air_passengers %>% 
  summarise_at(vars(starts_with("J")), mean)
```

Note the difference summarise_at and summarise_if:  


Now if we want to explore the `mutate_**()` functions family:
```{r}
# add columns for Jan/Feb/Mar with their median flights:
air_passengers <- air_passengers %>% 
  mutate_at(vars(one_of(month.abb[1:3])), funs(med = median))
```
